Forecasting the 2024 Election: Trump Dominates the South, Harris Leads in the Midwest

### 2024-10-22 paper/paper.qmd
committing text changes to paper.qmd

## Model set-up

The binary outcome variable $y_i$ is defined as 1 if Donald Trump is predicted to win state $i$, and 0 if Kamala Harris is predicted to win. The predictor variables are $\beta_i$, which represents the average percentage of polling support for Donald Trump in state $i$, and $\gamma_i$, the average percentage of support for Kamala Harris in state $i$. The relationship between these variables and the outcome is captured through a logistic link function.

@@

## Diagnostics {#sec-model-diagnostics}
Out-of-sample testing was conducted to test our model’s reliability in predicting potential presidential election outcomes. We train our model on a training set of our data and then test its performance on the test set of our data, we evaluate how well the model generalizes to new data that it was not trained on. This process helps reveal whether the model is overfitting to the training data or truly capturing the underlying patterns that apply more broadly. The test produces an accuracy score, RMSE, and a confusion matrix, each measuring different aspects of the model’s prediction. Accuracy is the proportion of correctly predicted outcomes compared to the total number of predictions made in the test set.

@@

# Visualizing the RMSE and accuracy in a plot with skinnier bars
ggplot(accuracy_rmse_df, aes(x = Metric, y = Value)) +
  geom_bar(stat = "identity", fill = "steelblue", width = 0.2) +  # Adjusted width for skinnier bars
  labs(y = "Value") +
  theme_light()

```
### Confusion Matrix
[@tbl-confusion-matrix] displays the model's predictive performance in distinguishing between state-level wins for Trump (coded as 1) and Harris (coded as 0). In this case, the model correctly predicted 12 Harris wins and 15 Trump wins (true positives and true negatives). However, there were 5 instances where the model incorrectly predicted a Trump win when Harris won (false positives), and 3 instances where it predicted a Harris win when Trump won (false negatives). This breakdown helps assess not only the overall accuracy but also how well the model distinguishes between close contests in different states, which is critical for predicting the outcome of the 2024 presidential election.

>>>>

## Model set-up
The binary outcome variable $y_i$ is defined as 1 if Donald Trump is predicted to win state $i$, and 0 if Kamala Harris is predicted to win. The predictor variables are $\beta_i$, which represents the average percentage of polling support for Donald Trump in state $i$, and $\gamma_i$, the average percentage of support for Kamala Harris in state $i$. The relationship between these variables and the outcome is captured through a logistic link function.

@@

## Diagnostics {#sec-model-diagnostics}
Out-of-sample testing was conducted to test our model’s reliability in predicting potential presidential election outcomes. We train our model on a training set of our data and then test its performance on the test set of our data, we evaluate how well the model generalizes to new data that it was not trained on. This process helps reveal whether the model is overfitting to the training data or truly capturing the underlying patterns that apply more broadly. The test produces an accuracy score, RMSE, and a confusion matrix, each measuring different aspects of the model’s prediction. Accuracy is the proportion of correctly predicted outcomes compared to the total number of predictions made in the test set. The testing was implemented using the following packages in [@citeR]: [@caret], [@rstanarm], [@ggplot2], [@dplyr].

@@

# Visualizing the RMSE and accuracy in a plot with skinnier bars
ggplot(accuracy_rmse_df, aes(x = Metric, y = Value)) +
  geom_bar(stat = "identity", fill = "steelblue", width = 0.2) +  # Adjusted width for skinnier bars
  labs(y = "Value") +
  theme_light()
```
### Confusion Matrix
[@tbl-confusion-matrix] displays the model's predictive performance in distinguishing between state-level wins for Trump (coded as 1) and Harris (coded as 0). In this case, the model correctly predicted 12 Harris wins and 15 Trump wins (true positives and true negatives). However, there were 5 instances where the model incorrectly predicted a Trump win when Harris won (false positives), and 3 instances where it predicted a Harris win when Trump won (false negatives). This breakdown helps assess not only the overall accuracy but also how well the model distinguishes between close contests in different states, which is critical for predicting the outcome of the 2024 presidential election.
